\papertitle

### 2014-03-20 final-report/paper.tex
Cool title
edits on case 1

For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragment}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state. The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.

@@

@@

The Myria web front-end server is written in Python and runs on Google App Engine\footnote{\url{https://developers.google.com/appengine/}}. \system's user-interface (UI) is embedded into Myria's web front-end. We build \system's UI using D3\cite{2011-d3} to deliver an interactive visualization experience to the user. D3 is a JavaScript framework for data visualization on the web.

\system's front-end gets data from the Myria server as JSON and CSV files. JSON is used to describe the physical queryplan. CSV files contain logs on query execution at the fragment and operator level as well as data exchange between workers at different stages of the query execution.

@@

The \fragment view comprises two charts: the utilization chart and the
operators chart. The utilization chart (Figure~\ref{fig:utilization_chart}) shows on
how many workers a fragment is running over time. The chart is used to
quickly reveal any patterns in the schedule, like tail latency, long periods of
idleness, etc. Execution skew is visible as humps.
We implemented \emph{focus+context}\cite{furnas1986generalized} in form of
a small brush (at the top of the figure) that allows the user to zoom-in and further
analyze the problematic area without loosing the context of the whole execution.

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/utilization_chart}
  \caption{The utilization chart.}
  \label{fig:utilization_chart}
\end{figure}

@@

\subsubsection{Overview over all fragments}
\label{sec:overview}

@@

The \overall shows small multiples of the utilization chart as described in the previous section.
In addition to identifying execution skew, the user can compare the fragments and find
correlations between the execution of different fragments. This view helps the user navigate to the
right fragment and is not intended to offer further inside into the causes or performance problems.

@@

\subsection{Case 1: Identifying Performance Bottlenecks using the \overall}

The \overall allows the user to quickly get an overview over the execution of all fragments
on all workers in the form of small multiples of the utilization charts as introduced in
Section~\ref{sec:fragment}. In Figure~\ref{fig:overview_skew} the user can see for example
that fragment~2 is very irregularly executed on different workers. If the graph is not a
single hill (as it is the case in Fragment 1, 3 and 4), workers are sleeping because they
either don't have enough data available to execute the operators or the output buffers
are full and data cannot be sent to the next fragment in the query graph.

The user has not identifies Fragment~2 to be a performance bottleneck and can use the other views
to further investigate its causes.

@@

This paper introduced \system, a visualization tool carefully tailored to help
database developers or users to quickly identify common distributed databases
performance bottlenecks, like tail latencies, data skew, etc. Our tool has been
successfully integrated with the Myria distributed database system, and has already
been used to identify several bugs, like poor physical query plan optimization,
tail latencies, poor data partitioning, which led us to believe the
visualization techniques we used are appropriate to help the user quickly
navigate to/find the cause of the performance problem. \system could be easily
plugged-in with another distributed database system, the only requirement being
a specific log data format. We explained how we collect logs in Myria and recommend
our approach to developers that are porting \system to a new DDBMS.

>>>>

For example, when querying how many workers are working on a certain fragment at a given time to produce the visualization described in Section~\ref{sec:fragments}, we used a query to select only events in the root operators (the operators that have no parent in the same fragment) for each fragment and a custom map function that carries state. The state is an integer that is incremented when the operator is called on a worker and decremented when the operator returns.

@@

We highly suggest that developers use the second approach when porting \system to a new DDBMS.

@@

The Myria web front-end server is written in Python and runs on Google App Engine\footnote{\url{https://developers.google.com/appengine/}}. \system's user-interface (UI) is embedded into Myria's web front-end. We build \system's UI using D3\cite{2011-d3} to deliver an interactive visualization experience to the user. D3 is a JavaScript framework for data visualization on the web. We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that were made possible using D3.

\system's front-end gets data from the Myria server using \system's back-end as JSON and CSV files. JSON objects are used to describe the physical queryplan. CSV files contain logs on query execution at the fragment and operator level as well as data exchange between workers at different stages of the query execution.

@@

The \fragment view comprises two charts: the utilization chart and the
operators chart. The utilization chart (Figure~\ref{fig:utilization_chart}) shows on
how many workers a fragment is running over time. The chart is used to
quickly reveal any patterns in the schedule, like tail latency, long periods of
idleness, etc. Execution skew is visible as humps.
We implemented a small brush (at the top of the figure) that
allows the user to zoom-in and further analyze the problematic area.

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/utilization_chart}
  \caption{The utilization chart. }
  \label{fig:utilization_chart}
\end{figure}

@@

\subsubsection{Overview over all fragments}
\label{sec:fragments}

@@

The \overall shows small multiples of the utilization chart described in the previous section.
In addition to identifying execution skew, the user can compare the fragments and find
correlations between the execution of different fragments. This view helps the user navigate between 
fragments and is not intended to offer deep insight into performance issues. Section~{\sec:case1} describes
a usecase for the \overall visualization.

@@

\subsection{Case 1: Identifying Performance Bottlenecks using the \overall}
\label{sec:case1}

The \overall view provides the user with a summary of the execution of all fragments
on all workers as described in
Section~\ref{sec:fragment}. In Figure~\ref{fig:overview_skew} the user can see for instance
that Fragment~2 is irregularly executed on multiple workers. 
If the graph is not monolithic (as it is the case in Fragment 1, 3 and 4), workers are sleeping because they
either (1) don't have enough data available to work on (data dependency problem) or (2) the destination worker's
input buffers are full thus forcing the source workers to stall.

@@

This paper introduced \system, a visualization tool carefully tailored to help
database developers or users to quickly identify common distributed databases
performance bottlenecks, like tail latencies, data skew, etc. Our tool has been
successfully integrated with the Myria distributed database system, and has already
been used to identify several bugs, like poor physical query plan optimization,
tail latencies, poor data partitioning, which led us to believe the
visualization techniques we used are appropriate to help the user quickly
navigate to/find the cause of the performance problem. \system could be easily
plugged-in with another distributed database system, the only requirement being
a specific log data format.
